---
title: "Quick Start Guide"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Get Started in 5 Minutes

This guide gets you productive with rpkgdev immediately. For complete documentation, see the main vignette.

## Installation

```{r}
# Install from GitHub
# remotes::install_github("blasbenito/rpkgdev")

# Load the package
library(rpkgdev)
```

## The Big Three

Learn these three functions and you're 80% there:

### 1. dev_load_all()

Load your package for interactive testing:

```{r}
dev_load_all()
```

**Use this:** After changing code, before testing functions in console.

### 2. dev_test()

Run your tests:

```{r}
dev_test()
```

**Use this:** After modifying functions, to verify nothing broke.

### 3. dev_document_and_check()

Update docs and run R CMD check:

```{r}
dev_document_and_check()
```

**Use this:** Before every commit. This is your quality gate.

## Your Daily Workflow

```{r}
# 1. Load package
dev_load_all()

# 2. Make code changes...

# 3. Test interactively in console...

# 4. Run automated tests
dev_test()

# 5. Before committing
dev_document_and_check()
```

That's it. Repeat this cycle throughout your day.

## Create a New Package

Want to start a new package with full infrastructure?

```{r}
deploy_template(
  path = "~/projects/mynewpkg",
  package_name = "mynewpkg"
)
```

This creates:
- Complete R package structure
- 30 development scripts in `dev/`
- Pre-commit hooks
- pkgdown website setup
- Test infrastructure
- Claude Code agents (optional)
- GitHub Actions CI/CD

The package opens automatically in RStudio, ready to use.

## Release to CRAN

Four functions, run in order:

```{r}
# Step 1: Prepare (update version, NEWS.md, spell check)
dev_release_prepare()

# Step 2: Local checks (must pass 0/0/0)
dev_release_check_local()

# Step 3: Remote checks (Windows, macOS, multi-platform)
dev_release_check_remote()

# Step 4: Submit to CRAN
dev_release_submit()
```

Wait for clean results from Step 3 before running Step 4.

## Beyond the Basics

### Check Spelling

Before release:

```{r}
dev_spell_check()
```

Add technical terms to `inst/WORDLIST`.

### Test Coverage

See what's tested:

```{r}
dev_test_coverage(html_report = TRUE)
```

Opens interactive report highlighting untested code in red.

### Build Website

Create pkgdown site:

```{r}
dev_pkgdown_build_site(preview = TRUE)
```

### Analyze Package

Get structure metrics:

```{r}
dev_analyze_structure()
```

Shows function counts, documentation coverage, test files.

## Common Parameters

All functions share consistent parameters:

- `pkg = "."` — Work on different package: `dev_test(pkg = "../otherpackage")`
- `quiet = FALSE` — Suppress output: `dev_test(quiet = TRUE)`

## Function Discovery

Type prefix + TAB to discover related functions:

- `dev_` + TAB → All development functions
- `dev_check_` + TAB → All checking functions
- `dev_release_` + TAB → Release workflow (4 steps)
- `dev_analyze_` + TAB → Analysis functions

## Getting Help

```{r}
# See all functions
help(package = "rpkgdev")

# Function help
?dev_test
?deploy_template

# Read main vignette
vignette("rpkgdev")
```

## Next Steps

1. **Try the Big Three** on your current package
2. **Read dev/README.md** in deployed template for detailed script documentation
3. **Explore function families** using autocomplete
4. **Visit the website:** https://github.com/blasbenito/rpkgdev

That's all you need to get started. The rest you'll pick up as you go.
